### 2012-03-11 search.tex
finally finished an edit of all of section 3 (with a few holes)

search :: SearchStrategy -> Scorer -> [Seed]
       -> (Scored Placement, [Score])
search strat score (s0:seeds) = runFrom seeds firstGen [] 0
 where
  firstGen = map score $ gen0 strat s0
  runFrom :: [Seed] -> [Scored Placement] -> [Score]
          -> Age -> (Scored Placement, [Score])
  runFrom (s1:s2:seeds) oldPop oldHist age =
    let trialPop  = nextGen strat s1 score oldPop
        trialHist = (scoreOf $ minimum newPop) : oldHist
        (newPop, newHist) =
          if accept strat s2 newHist age then
            (trialPop, trialHist)
          else
            (oldPop, oldHist)
    in  if quit strat newHist age then
          (minimum newPop, newHist)
        else
          runFrom seeds newPop newHist (age + 1)