### 2012-05-08 icfp_reviewer_response.txt
reviewer response updated

We thank the reviewers for their insigful comments, and if our manuscript is 
accepted, we will incorporate changes that reflect these comments. We detail 
our responses to the reviewers below:

Reviewer A writes:

-- More discussion of the comparison between previous C++ software in 
-- that authors' research group and the new Haskell software would be 
-- much more interesting. In particular, the C++ tools described in 
-- section 4 seem significantly more complex than MRFy. Would Haskell 
-- scale to these problems just as well?

The C++ tools described are actually no more complex in terms of the algorithms 
they implement, though they comprise many more lines of code. We will clarify 
this language, and better contrast them.

-- Additionally, the performance discussion is so brief as to be almost 
-- uniterpretable. For example, it seems that MRFy parses protein 
-- structure (in what format?) in constant time, whereas SMURF seems to 
-- take longer on more complex proteins. Are the two tools implementing 
-- the same algorithm? Why is Haskell slower by a factor of 3 on the 
-- small structures? The discussion on parallelism is even briefer -- 
-- it's great to keep 48 cores warm, but what kind of parallel speedup is 
-- achieved?

We agree, and will provide more comprehensive quantitative results. In point of 
fact, MRFy is not implementing the same algorithm as SMURF, and we will better 
clarify this. We will also correct our wording to indicate that it's not 
*parsing* protein structures that we're talking about, but *computing* the 
optimal parse of a protein sequence onto a Markov random field. The complexity 
of the protein structures used to build the Markov random field is at issue 
here, and we will make this clearer. We will provide a better analysis of 
parallel performance, as well. In addition, while SMURF and MRFy differ 
algorithmically, we can and will provide a quantitative comparison of the 
run-time performance of the algorithm they share in common: Viterbi's algorithm.

-- The discussion of Viterbi's equations is extremely hard to follow for 
-- the uninitiated. This is especially true since the level of detail 
-- presented doesn't seem necessary for the larger points of the paper.

We agree, and given the request for more quantitative results, we will try to 
clarify the Viterbi discussion while at the same time going into less detail.

-- Is MRFy available? Can we try it out?

We plan to make MRFy publicly available by the time ICFP opens (and by which 
time we will have also published scientific results based on MRFy). We can make 
MRFy available for review if desired, but the primary investigator (Daniels' 
PhD advisor) is reluctant to release source code prior to publication of 
results.

-- I'm not sure that other structural biologists would agree that 
-- searching for homologous proteins is the "best known method of 
-- discovering" protein function.

We will clarify this to indicate that searching for homologous proteins is a 
well known and cost-effective approach that works in the absence of genetic 
interaction networks, given that protein sequence information is widely 
available. We will cite Pandey et al. 2007, "Computational Approaches for 
Protein Function Prediction: A Survey"

-- Why does the definition of HMMState on page 2 contain no data, just 
-- tags?

We realize that our use of "state" is misleading; our representation uses this 
HMMState data type as a tag, which along with node index and sequence residue 
index, allows us to index into our memoization table. We will change the data 
type to "StateLabel" and clarify this in the manuscript.

Reviewer B writes:

-- There is a report of positive experience using parallelism in Haskell, which is 
-- striking in its brevity. You could easily miss it! Apparently, at least for 
-- trivially parallel problems such as this, parallelism with GHC just works and 
-- there was nothing more to say. I'd like to see a bit more made of this result, 
-- since as a community we're often asked for evidence to support our hypothesis 
-- that functional programming is good for parallelism. Perhaps give parallelism a 
-- separate subsection and include a few numbers?

We plan to break parallelsm out into its own subsection, and include a 
quantitative analysis of scaling on several multi-core systems.

-- 3.1 
--    "data TProb" - why not use a newtype?

This is a historical artifact; TProb used to contain multiple fields. We will 
fix it.

-- 3.2 
--    - the code for "search" is a bit compressed - consider adding some line
--      breaks and blank lines to make it more readable

We will give "search" a bit more space.

-- 3.3 
--    - it sounds like your solution for vee' might be relying on lazy 
--      evaluation rather than properties of GHC's inliner, but it's hard 
--      to tell without more details.

We don't believe this to be the case, but it seems like a worthwhile and simple 
experiment. We will attempt to add strictness to vee' and report whether or not 
there is any impact.

-- 3.4 
--    - since version 7.4.1, GHC's profiler now annotates nested 
--      functions automatically.

This is terrific news! We plan to try GHC 7.4.1 once we resolve some library 
dependencies that are currently problematic. We will update the text to reflect 
this new support.

--   - it's not clear why the profiler could only be used on very small 
--     test cases: is it speed, memory use, or something else? If speed, 
--     does it help to use fewer annotations?

It was definitely runtime speed that was the problem here, for any test case 
large enough to be interesting, even if few annotations. We will state this 
more clearly and provide a more thorough evaluation of our profiling 
experiences.

-- 4. 
--    "working with HMMER codebase" -> "working with the HMMER codebase"

We will correct this.

Reviewer C writes:

-- The main thing I'm missing is more discussion of the effect of laziness -- in 
-- general and on the specific points mentioned. Was laziness useful for the 
-- algorithms described? Was it perhaps even instrumental? Or was it rather a 
-- hurdle (in particular with respect to the issues of debugging, profiling and 
-- performance)? Perhaps the authors can try to make some informed guess whether, 
-- with everything else being equal, a strict language might or might not have 
-- worked better for their problem domain.

We will include more discussion of laziness. We used laziness in two essential 
places, though one of them was not preserved: our original bottom-up 
dynamic-programming approach to Viterbi, and our use of a lazy infinite list of 
random numbers in our search implementation. We will discuss this in more 
detail.

In general, though, we don't particularly feel that laziness had a significant
bearing on our experience; other than one bug which we thought was due to
laziness, but actually turned out to be a mis-implemented typeclass derivation,
laziness was neither an impediment nor instrumental to our success.

-- Sec 3.1: In the type definition for Scored, what is type Score?

We will add this to section 3.1; as it happens it is "newtype Score = Score 
Double"

-- Does the code for vee' omit some cases, or what's happening for other 
-- HMMStates? Also, what is eProb?

Yes, this is only the Match case; we omitted the other recursive and base cases 
in the interests of brevity. We will enumerate the cases we've omitted, and 
refer the reader to Figure 3, but it would require too much space to include 
all of the code.

-- Sec 3.2: "Our implementation of hidden Markov models and Viterbi's algorithm 
-- duplicate existing functionality." I didn't get this. What functionality?

We will clarify this; we meant that we were reimplmenting some algorithms 
(namely, Viterbi) that have already been implemented (in C++) in SMURF and 
HMMER. The point of this contrast is to acknowledge that had we simply extended 
the SMURF code base to implement our new algorithm, we would not have had to 
reimplement Viterbi, for better or worse.

-- In the type definition for Scorer, what is type Placement?

In this case, [Int], but our real code is polymorphic. We ask the reviewers and 
committee: should we show the polymorphic version instead, though it may be 
less clear how it relates to this particular problem domain?

-- You say the use of higher-order functions was useful, e.g. for abstracting over 
-- search strategies. But isn't your type SearchStrategy effectively defining an 
-- object type? That is, wouldn't this have been as convenient with conventional 
-- object-oriented techniques?

We don't believe this would have been as convenient with conventional 
object-oriented techniques, as we are closing over a great deal of environment 
information (command-line parameters, some of which are search-strategy 
specific) which would have required instantiating various types of 
search-strategy-specific objects.

-- Sec 4: Typo: "Matt uses such data structures such as..." - double "such"

Indeed; we will fix this.

Reviewer D writes:

-- Perhaps the most useful feature of Haskell that is used here is function 
-- memoization which comes naturally to dynamic programming and is cumbersome in 
-- other programming languages save for the efficiency of the implementation of 
-- memoization in Haskell itself. It would have been useful to compare this to 
-- implementing table lookups in other programming languages.

We can provide a contrast to our experience implementing dynamic programming in 
C++ (unsurprisingly, there is a great deal more bookkeeping to be done). We do 
not feel we have the experience to provide a comparison over a wide range of 
programming languages.

-- While in the introduction the authors state that "considerations of performance 
-- are paramount" no numerical comparison is presented in the paper about 
-- performance of their implementations compared to others written in C or C++. 
-- Also, no viable way of comparing "convenience, readability, and productivity" 
-- to other implementations is presented in a quantitative manner that can help 
-- attract would-be Haskell programmers except for section on the use of 
-- memoization.

We promise a quantitative comparison should this manuscript be accepted. This 
comparison will largely have to be limited to code size and development time, 
but we can and will provide a quantitative comparison of the Viterbi algorithm 
run-time performance between SMURF and MRFy. However, the overall MRFy 
algorithm differs from SMURF, so overall run-time differences cannot be 
attributed to language choices as opposed to algorithmic differences.

-- Can the Authors identify other problems in computational biology, other than 
-- protein amino acid sequence comparison, where their experience presented in the 
-- paper would apply and be useful? I doubt it. For example in Section 5, "What 
-- can you learn from our experience?" The Authors state "If you are a 
-- computational biologist, you can be productive in Haskell, and you don’t need 
-- extensive experience." I would ask them to name one more problem in 
-- computational biology that would make use of their experience.

In fact, we can name several other problems we can (and may) address based on 
this experience. For example, we have applied graph-theoretic algorithms to the 
problems of gene functional annotation given a genetic interaction graph, as 
well as the problem of identifying compensatory genetic pathways 
("Between-pathway models") in the Yeast genome. We have written some Haskell 
code for the former problem (very much a work in progress), and while we 
approached the latter problem using Python, we feel that Haskell would have 
suited nicely. We will expand upon this discussion in the paper.

-- As stated earlier, identifying homologous proteins is a very important problem, 
-- but it is misleading to state that there are no other problems in computational 
-- biology that are important, the filed so very vast and problems so numerous 
-- which makes the title of the paper misleading. If the Conference Committee 
-- decides to accept the paper, I would urge them to ask the Authors to change the 
-- title of the paper.

If accepted, we will choose a more appropriately narrow title. It was not our 
intent to suggest that protein remote homology detection is the only field 
within computational biology, but rather that many computational biologists are 
unfamiliar with Haskell, and most computational biology software is written, as 
we suggest in the introduction, in C++, Java, or Python.

-- Also, In making a decision to accept or reject the paper, the Conference 
-- Committee should compare their decision with that of accepting or rejecting a 
-- paper that describes implementation of a dynamic programming algorithm in 
-- Hasekell. The phrase "computational biology" should not be a factor in their 
-- decision; the same methods presented in the paper could have been used in 
-- implementing an algorithm for comparing sequences of strings over a finite 
-- alphabet without change, regardless of the fact that in this case the alphabet 
-- encodes names of amino acids.

Clearly, we have not adequately explained the difference between our 
implementation (which combines dynamic programming, stochastic search, and a 
relatively new graphical model for Markov random fields, which are not the same 
as hidden Markov models) and past work from our group. We intend our experience 
report to be an exemplar of what computational biologists may experience should 
they choose to implement novel, performance-intensive algorithms in Haskell 
(or, perhaps, other functional languages). We will endeavor to make this goal 
clearer should our paper be accepted.

We would also like to better explain that MRFy represents a novel algorithm for 
solving Markov random fields, but whose components include implementations of 
the standard techniques of dynamic programming and stochastic search. We hope 
that some readers not yet familiar with Haskell will find this higher-order 
stochastic search implementation useful as well.