### 2012-06-20 everygen.tex
damnfino what's going on, but these are my changes

everyPt :: SearchGen pt r -> Age -> Scored pt
         -> Rand r [Aged (Utility (Scored pt))]
everyPt ss age startPt = do
  successors <- mapM (nextPt ss) (repeat startPt)
  tagged <- zipWithM agedUtility successors [succ age..]
  let (useless, Aged (Useful newPt) newAge : _) =
                        span (isUseless . unAged) tagged
  nextPts <- everyPt ss newAge newPt
  return $ Aged (Useful startPt) age : useless ++ nextPts

  where agedUtility pt age =
           utility ss move >>= \u -> return $ Aged u age
         where move = Move { older = startPt, younger = pt
                           , youngerAge = age }