### 2012-06-20 orig-search.tex

search :: forall placement 
        . SearchStrategy placement 
       -> Scorer placement 
       -> [Seed]
       -> (Scored placement, History placement)
search strat scorer (s0:seeds) = runFrom seeds firstGen (History []) 0
 where
  firstGen = map scorer $ gen0 strat s0
  runFrom :: [Seed] -> [Scored placement] -> History placement
          -> Age -> (Scored placement, History placement)
  runFrom (s1:s2:seeeds) oldPop oldHist age =
    let trialPop  = nextGen strat s1 scorer oldPop
        trialHist = (winner, age) `hcons` oldHist
        winner = minimum trialPop
        (newPop, newHist) =
          if accept strat s2 trialHist age then
            (trialPop, trialHist)
          else
            (oldPop, oldHist)
    in  if quit strat newHist age then -- TODO quit must change: consider best-ever, convergence
          (fst $ minimum (unHistory newHist), newHist) 
        else
          runFrom seeds newPop newHist (age + 1)