### 2012-06-20 orig-strategy.tex

type Age  = Int -- number of generations explored
type Seed = Int -- source of stochastic variation
data History placement = History { unHistory :: [(Scored placement, Age)] }
  deriving (Show, Eq, Ord)
hcons :: (Scored placement, Age) -> History placement -> History placement
hcons a (History as) = History (a:as)
emptyHistory :: History a
emptyHistory = History []

-- is there a better name for seed?
data SearchStrategy placement = 
 SS { gen0    :: Seed -> [placement]
    , nextGen :: Seed -> Scorer placement
              -> [Scored placement] -> [Scored placement]
    , accept  :: Seed -> History placement -> Age -> Bool
    , quit    ::         History placement -> Age -> Bool
    }