### 2012-06-20 search.tex

search :: SearchStrategy pt r -> Rand r (History pt)
search (SS strat test) =
  fmap test . everyPt strat 0 =<< pt0 strat