### 2012-06-28 everygen.tex
wordsmithing, death of age, revised search

everyPt :: RandomGen r
        => SearchGen pt r -> CCost -> Scored pt
        -> Rand r [CCosted (Utility (Scored pt))]
everyPt sg age startPt = do
  successors <- mapM (nextPt sg) (repeat startPt)
  tagged <- zipWithM agedUtility successors [succ age..]
  let (useless, CCosted (Useful newPt) newCCost : _) =
                        span (isUseless . unCCosted) tagged
  nextPts <- everyPt sg newCCost newPt
  return $ CCosted (Useful startPt) age : useless ++ nextPts

  where agedUtility pt age =
           utility sg move >>= \u -> return $ CCosted u age
         where move = Move { older = startPt, younger = pt
                           , youngerCCost = age }