### 2012-06-29 search.tex
more on search

search :: RandomGen r => SearchGen pt r -> SearchStop pt
       -> Rand r (History pt)
search strat test =
  return . test =<< everyPt strat 0 =<< pt0 strat