### 2012-07-01 search.tex

search :: RandomGen r
       => SearchStrategy pt r -> Rand r (History pt)
search (SS strat test) =
  fmap test . everyPt strat 0 =<< pt0 strat