### 2013-04-28 jfpnotes.tex
jfp draft with notes

\subsection{Introduction}

A preliminary version of this paper was presented as an \emph{Experience
Report} at ICFP 2012. In this revised paper, we add to our original work
the results of two different attempts to improve and deepen the
application of functional programming to the detection of remote
homologies in proteins:

\begin{itemize}
\item
  To detect the likelihood that a protein is drawn from a family
  represented by a given Hidden Markov Model, we implemented Viterbi's
  algorithm in Haskell. The algorithm uses dynamic programming, which we
  implemented by memoizing a recursive function.

  Our original Haskell code ran 7 to 8 times slower than C++ code built
  in the same research group for the same purpose. Because MRFy exploits
  parallelism well, we could afford to give up this factor of 7 or 8,
  but we felt this number was awfully high. In Section XXXX we report
  our efforts to speed up a native Haskell implementation.
\item
  Our code for Viterbi's algorithm is carefully crafted to reflect the
  equations usually used in textbooks to describe the algorithm. But an
  experienced functional programmer might find our original code
  distasteful:

  \begin{itemize}
  \item
    The central data structure is an array, and it is used in a way that
    is more typical of FORTRAN than of any functional language. (Lots of
    index arithmetic, no wholemeal programming.)
  \item
    There is a good deal of fiddling with indices at the boundaries.
  \item
    The elegant structure of the Plan 7 Hidden Markov Model is
    \emph{not} reflected in the static types used in the Haskell code.
  \item
    There are an alarming number of cases.
  \end{itemize}

  In Section YYYY, we report on our efforts to make the data structures
  and algorithms ``more functional.'' We refactor our original code and
  assess it on three dimensions:

  \begin{itemize}
  \item
    How much is it still manifestly an implementation of Viterbi's
    equations?
  \item
    How well do the static types model the structure of the problem?
  \item
    How do we feel about the code?
  \item
    How does making the code ``more functional'' affect performance?
  \end{itemize}
\end{itemize}

\subsection{XXXX: Experience trying to make Haskell competitive with
C++}

We asked, according to our relatively naive understanding of what
factors affect performance, how Haskell code might differ from C++ code.
We came up with these ideas:

\begin{enumerate}[1.]
\item
  In C and C++, record structures compose with no additional cost in
  allocations or indirections. That is, a record of records is allocated
  in a single block of memory, and a member of a member can be addressed
  with only one address-arithmetic operation---there is no
  ``pointer-chasing''.

  In Haskell, because the language is polymorphic, we believe that a
  typical value must be represented in a single machine word. If the
  value is too big to fit in a machine word, it must be allocated on the
  heap and represented by a pointer. Such a representation is called
  ``boxed''.

  Because Haskell is lazy, even individual members of a record may be
  either evaluated or unevaluated, and may therefore be boxed.

  We set out to make the \emph{machine-level} representation of our
  Haskell data structures more like the machine-level representation
  that a C programmer would use. We were able to discover two mechanism:

  \begin{itemize}
  \item
    A \emph{strictness annotation} can force the evaluation of a part of
    a record and therefore prevent it from being boxed in a thunk.
  \item
    The \texttt{UNPACK} pragma: where do we learn about it (aside from
    Johan Tibell), what do we cite, how is it known how to use it?
  \item
    Does GHC provide an option that enables us to look at some kind of
    GHC IR and know if we have succeeded?
  \end{itemize}
\item
  C and C++ use pointer arithmetic and unsafe array access.

  \begin{itemize}
  \item
    What happens if we change the record of transition probabilities to
    an array indexed by pair of states?

    What should the index type be?

    Can we look at GHC output (maybe \texttt{-ddump-simpl}) and figure
    out whether we are getting a flat two-dimensional array vs an array
    of arrays?

    Is there a one-dimensional array that works?

    Since we are indexing using every element of an enumeration type,
    what is a way we can introduce unsafe indexing that we believe is
    safe according to the type system? Can we convince the compiler?
  \end{itemize}
\end{enumerate}

\subsection{YYYY: Experience making things ``more functional''}

After the experience of presenting our work at ICFP and meeting other
functional programmers there, we became aware that our central data
structure is not what a functional programmer might expect to see. We
were curious to see what effect it might have to change this data
structure. {[}questions from above{]}?

Diagramatically, a Hidden Markov Model looks like a Begin node followed
by zero (one?) or more ordinary nodes followed by an End node. But in
our original code, all nodes are represented in the same way, and the
distinct properties of Begin and End nodes are present only in the code,
not in the types. (As new Haskell programmers, we tended to choose
representations similar to the C representations we had used before, and
because C does not provide discriminated-union type, the responsibility
for knowing which node we're looking at \emph{always} resides in the
code. Our groups C code does not even use C unions (\textbf{check
this}); there would be little benefit.)

\begin{itemize}
\item
  Because every node has to be capable of representing a Begin node,
  every node stores two probabilities used only for transitions out of
  Begin nodes, even though in all but one of the nodes, these
  probabilities go unused.
\item
  (Mumbo jumbo about chopping the node sequence into pieces, so that any
  node might become a Begin node after beta strands are placed. Again,
  the different interpretations of the ``same'' node are not explicit in
  the type system.)
\end{itemize}

Experiments to be carried out:

\begin{enumerate}[1.]
\item
  Refactor the basic representation to eliminate the unused transition
  probabilities.
\item
  Introduce an explicit Begin node with its own custom representation.
\item
  Create a \texttt{Model} type with the form
  \texttt{Begin \{Middle\} End}. Inspired by REPA-style indexing tricks
  (also seen elsewhere), represent the middle sequence as a function of
  type \texttt{Int -\textgreater{}     Middle}. Implement a slicing
  function
  \texttt{Array Node -\textgreater{} Interval -\textgreater{}     Model}.
\item
  Rewrite Viterbi with the new representation.
\end{enumerate}

N.B. The elimination of unused transition probabilities is
\emph{incompatible} with the change in representation to use a
two-dimensional array. These experiments will have to be handled
carefully; we may need to build all four variants from

\begin{verbatim}
{ one node type, three node types } * { probability record, probability array }
\end{verbatim}

A more radical change in the representation is to use the functional
programmer's mainstay: a list instead of an array. This means lists of
nodes as well as lists of residues. This change was motivated by two
observations:

\begin{itemize}
\item
  Our Viterbi code was doing an awful lot of array indexing and index
  arithmetic. This felt more like FORTRAN than like Haskell.
\item
  There were a \emph{lot} of special cases around indices like 0 and -1.
  {[}Some of these cases arose because we didn't have a distinct Begin
  node, so we'll have to see what the code looks like when a Begin node
  is introduced.{]}
\end{itemize}

Changing to lists had several effects:

\begin{itemize}
\item
  The algorithm suddenly came into focus as something familiar to
  functional programmers. Instead of being something blindly translated
  from equations in a textbook, it became (probabilistic) minimum edit
  distance. We thought this was pretty cool.
\item
  It's no longer so obvious how our code relates to Viterbi's algorithm
  as explained in textbooks. Instead, we are relying on a deeper
  understanding of the underlying problem. Perhaps good for us, but
  maybe not so good for communicating with, e.g., new graduate students
  in the group. The jury is still out here.
\item
  Memoization is \emph{much} hairier. The junior members of the team
  might not have been able to pull this off on their own. Lots of
  QuickCheck action, and significant uncertainty.
\item
  We've knowingly introduced new pointer indirections, with who knows
  what effect on locality or on memory traffic. How did this affect
  performance? {[}This problem militates toward testing on multiple
  hardware platforms, including NR's ancient AMD chip before it goes
  away.{]}
\end{itemize}

\subsection{Benchmarking}

We created several benchmarks without beta strands so that MRFy
\emph{only} executes a single pass of Viterbi without executing a
stochastic search. For every benchmark, we tested a change in either the
implementation of \texttt{viterbi} or a change in the representation. We
report figures relative to a baseline, which was established to account
for IO and parsing.

\subsection{Naming the parts of the design space, Take One}

\{Since there are many changes, my idea here is to give names to each of
the changes. Ideally, they'd correspond to a branch name. I have not
thought this through.\}

\begin{itemize}
\item
  \texttt{special-directives} - \{I don't know what this is. These were
  suggested by John Tibell and apparently already added?\}
\item
  \texttt{indirection-removal} - \{I don't know what this is. See
  \#11.\}
\item
  \texttt{state-arrays} - In \texttt{viterbi}, instead of using case
  analysis on HMM states, we use arrays indexed by HMM states.
\item
  \texttt{unsafe-arrays} - Use unsafe array ranges for memoization.
\item
  \texttt{separate-begin} - The ``begin'' HMM node requires special case
  analysis, so we therefore explicitly separate it from all other HMM
  nodes.
\item
  \texttt{static-trans} - Static encoding of HMM transition
  probabilities. \{??? See issue \#2.\}
\item
  \texttt{HMM-list} - Represent an HMM as a \emph{list} of nodes and
  memoize on its prefix.
\end{itemize}

The number of all possible tests given these changes is quite large, so
we tested each change individually and tested some interesting
combinations of changes: \{I don't think we know yet.\}

\{We talked about a commutative diagram in issue \#8. But it seems like
we don't need that for all tests---maybe only the \texttt{static-trans}
and \texttt{HMM-list} tests? What about \texttt{separate-begin}? With
\texttt{separate-begin}, what if we constructed the special node within
\texttt{viterbi}?\}

\subsection{Naming the parts of the design space, Take Two}

The design space is combinatorial. I suggest we name the individual
elements:

\begin{itemize}
\item
  Records of records: nested, flat

  \begin{itemize}
  \item
    relates to special directives and indirection removal?
  \end{itemize}
\item
  Transition tables: named fields, single array, array of arrays
\item
  Transition array indexing: safe, unsafe
\item
  Viterbi decreases: lists, array indices
\item
  Node array indexing: safe, unsafe, not used in inner loop
\item
  Residue array indexing: safe, unsafe, not used in inner loop
\end{itemize}

Some of these are independent and some are connected. For example, if
Viterbi decreases array indices, then that indexing must be safe or
unsafe. If Viterbi decreases lists, then array indexing is not used in
the inner loop.