### 2013-05-27 hoviterbi.tex
code updates

vee' stateRight j i =
  internal [ child score state (vee'' state pj pi)
           | state <- preceders stateRight
           , let score = transition node state stateRight
                         + emission node state aa
           ]
  where pi = prevres  stateRight i
        pj = prevnode stateRight j
        node = middle pj
        aa = rs `rat` i
       vee' stateRight nodes@(node:ntail) aas =
         internal [ child score state (vee'' state (nextNodes stateRight)
                                         (nextAAs   stateRight))
        | state <- preceders stateRight
        , hasAA stateRight aas
        , let score = transition node state stateRight
                      + emission node state (head aas)
        ]
         where -- Del does not consume a residue
     -- Ins does not consume a node
     nextNodes Ins = nodes
     nextNodes _   = ntail
     nextAAs Del   = aas
     nextAAs _     = tail aas
     hasAA Del _   = True
     hasAA _  aas  = not (null aas)