### 2013-05-27 everygen.tex
paper changes

everyPt :: RandomGen r
        => SearchGen pt r -> CCosted (Scored pt)
        -> Rand r [Utility (CCosted (Scored pt))]
everyPt sg (CCosted cost startPt) = do
  successors <- zipWithM costedUtility [succ cost..] =<<
                mapM (nextPt sg) (repeat startPt)
  let (useless, Useful pt : _) = span isUseless successors
  (++) (Useful (CCosted cost startPt) : useless) <$> everyPt sg pt
 where
   costedUtility cost pt = return . fmap (CCosted cost) =<< utility sg move
     where move = Move { older = startPt, younger = pt, youngerCCost = cost }