### 2013-05-27 hov-prevs.tex

prevs preceders stateHat fj fi =
 internal [ edge score state (vee'' state pj pi)
          | state <- preceders
          , let pi = fi state
          , pj >= 0, pi >= 0
          , let score = transition (node pj) state stateHat
                        + emission (node pj) state (residue pi)
          ]
 where pj = fj stateHat

predUnless :: forall a . Enum a => a -> StateLabel -> StateLabel -> a
predUnless n don't_move s = if s == don't_move then n else pred n