### 2013-05-27 search.tex

search :: RandomGen r
       => SearchGen pt r -> SearchStop pt -> Rand r (History pt)
search sg test = return . test =<< everyPt sg . CCosted 0 =<< pt0 sg