### 2013-05-28 hov-prevs.tex
more code snippets

       prevs predecessors stateHat fj fi =
        internal [ edge score state (vee'' state pj pi)
                 | state <- predecessors
                 , let pi = fi state
                 , pj >= 0, pi >= 0
                 , let score = transition (node pj) state stateHat
                               + emission (node pj) state (residue pi)
                 ]
        where pj = fj stateHat
prevs :: [StateLabel] -> StateLabel
      -> (StateLabel -> NodeIndex) -> (StateLabel -> ResidueCount) -> a