### 2013-05-28 hov4.tex

vee' :: StateLabel -> NodeIndex -> ResidueCount -> a
vee' Del (NI 1) (RC 0) = leaf (transition (node 0) Mat Del)
vee' Ins (NI 0) (RC 0) = leaf (transition (node 0) Mat Ins)
vee' Mat (NI 1) (RC 0) = leaf (transition (node 0) Mat Mat)
vee' stateHat j i = 
  internal [ edge score state (vee'' state pj pi)
           | state <- preceders stateHat
           , let pi = predUnless i Del state
           , pj >= 0, pi >= 0
           , let score = transition (node pj) state stateHat
                         + emission (node pj) state (residue pi)
           ]
  where pj = predUnless j Ins stateHat