### 2013-05-28 acm-para.tex
eps of search graph 5. two not very good paragraphs on benchmark and acm submission.

A paper for the computational biology community, based on MRFy, has been
submitted to the ACM Conference on Bioinformatics, Computational Biology, and 
Biomedical Informatics.
This paper explores the effectiveness of various search strategies, and also
compares MRFy's accuracy at the task of remote homology detection to several
contemporary approaches.
MRFy outperforms the other software tested.
On these remote homology detection tasks, MRFy spends virtually all of its
time in the Viterbi algorithm, so improvements to this implementation will
measurably improve MRFy's effectiveness.