### 2013-05-28 medium-para.tex

We primarily benchmarked MRFy's run-time performance using a test case that
mimics real-world use of MRFy, but focuses on the implementation of the Viterbi
algorithm.
This ``medium-sized'' benchmark consists of computing the Viterbi score for
each of a set of protein sequences against each of a set of hidden Markov model
templates.
The medium-sized benchmark comprises 56 protein sequences from the deep-sea 
bacterium \emph{Thermotoga maritima},
ranging in length from 80 to approximately 1200 residues, and 9 hidden Markov
models, each built from a multiple structure alignment of the proteins in
a $\beta$-structural protein superfamily.
A single invocation of MRFy results from each hidden Markov model in the 
benchmark, and that invocation aligns each of the 56 protein sequences in the 
input file to that model. Thus, time spent in I/O and parsing of the model
is amortized over all protein sequences.