### 2013-05-29 localsearch-para.tex
2 paras about local search. Missing citation; I've emailed Richard to ask how he'd like it cited.

Richard Senington proposed and prototyped a ``local search'' model of
stochastic search, using his \texttt{local-search} library~\cite{}.
This library did not fit well with our stochastic search abstraction, and
thus we wrote an alternative MRFy implementation using \texttt{local-search}
as well as the majority of the MRFy modules.
While reconciling these two codebases will constitute a tedious task, and we are
not happy with this current state of MRFy, the results were terrific.
The local search approach outperformed random hill-climbing, simulated
annealing, and the genetic algorithm.

Our hasty modifications to MRFy, in order to incorporate local search, led to
good results quickly.
We did not need to change any interfaces in the various MRFy modules, though in
the interim, we have a separate \texttt{main} for local search.
This stands in contrast to our experiences modifying our C and C++-based
software, which inevitably led to headaches.